Supplementary Information

### 2020-11-16 doc/manuscript/MEE_manuscript.Rmd
climwin figures in SI

#SI figure numbers
traditional_comparison = "S1"

#SI figure captions
traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals."

```

@@

**Our process picked out similar climate drivers to what would be obtained via traditional methods (Figs. `r SCBI_example`, **S2**; Table `r qualitative_results_comparison`; Appendix `r traditional_comparison`), but with the advantage that *climwin* allowed objective selection of the strongest climate drivers and the time windows over which they were most influential.**
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, each of which was identified by climwin as the top temperature-related driver *at four of the eight sites*. $T_{min}$ was the top driver at only *one site (BCI)*, and $T_{mean}$ was never identified as the top variable within the temperature group (Fig. `r clim_sens`).
Within the precipitation group, precipitation amount ($PPT$) was identified as the top variable most frequently (n=*6 of 9* sites), but it was not uncommon that it was surpassed by precipitation frequency ($PDF$; n=*3 of 8* sites). 
Optimal time windows often coincided with a site's peak growing season (n= *# of 10* for temperature variables, *# of 10* for precipitation variables), but exceptions were common. 
At the *5 lowest latitude* sites (BCI, HKK, LT, CB, and SCBI), the optimal window for precipitation variables spanned $\ge$ 8 months, ending during the peak growing months of the year of ring formation. 
At the *3 highest latitude sites* (HF, ZOF, and SC), the optimal window for precipitation variables was a short ($\le$ 3 months) window during the previous growing season.
Optimal windows for temperature variables tended to be shorter, the longest being a *6*-month period during the summer (wet season) at HKK. 
At two of the higher-latitude temperate sites (HF and Žofín), temperatures were most influential during late winter/ early spring.
There were also a few instances where previous growing season conditions had the strongest influence.

@@

**Across the three metrics of growth, the "landscape" of climate effects over various time windows was generally similar, but the optimal time window or even the top climate variable sometimes differed (Figs. **S2**-**S4**).**
Specifically, $RW$, $BAI$, and $\Delta AGB$ consistently exhibited similar strength of correlation and direction of response to climate variables within the temperature and precipitation variable groups. 
In some cases (n= # of #), both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. **S2**). 
In #cases, *climwin* identified the same climate variable but different (sometimes overlapping) time windows. 
In *2* cases (precipitation variable group at LT, temperature variable group at HKK), *climwin* identified different climate variables, but identical or overlapping time windows (e.g., Fig. **S3**).
Finally, in *2* cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at BCI, precipitation variable group at HF; Figs. `r clim_sens`, **S5, S11**), *climwin* identified different climate variables and different time windows (e.g., Fig. **S4**). 
Henceforth, we focus on the climate drivers identified when $RW$ was the growth metric and for the full set of cores (i.e., including those for which $DBH$ could not be reconstructed.)

>>>>

#SI figure numbers
all_traditional_comparisons = "S1-S4"
SCBI_traditional_comparison = "S1"
Zofin_traditional_comparison = "S2"
CB_traditional_comparison = "S3"
SC_traditional_comparison = "S4"

all_climwin = "S5-S7"
climwin_sameV_sameW = "S5"
climwin_sameV_diffW = "S6"
climwin_diffV_diffW = "S8"

#SI figure captions
SCBI_traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals: LITU at SCBI."
Zofin_traditional_comparison_legend <- "Figure S2. Comparison of our approach with traditional methods of identifying climate signals: ABAL at Zofin."
CB_traditional_comparison_legend <- "Figure S3. Comparison of our approach with traditional methods of identifying climate signals: PSME at Cedar Breaks."
SC_traditional_comparison_legend <- "Figure S4. Comparison of our approach with traditional methods of identifying climate signals: PIMA at Scotty Creek."

climwin_sameV_sameW_legend = "Figure S5. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S6. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S7. (TMP/TMN at BCI)"
```

@@

**Our process picked out similar climate drivers to what would be obtained via traditional methods (Figs. `r SCBI_example`, `r all_traditional_comparisons`; Table `r qualitative_results_comparison`; Appendix `r traditional_comparison`), but with the advantage that *climwin* allowed automated, objective selection of the strongest climate drivers and the time windows over which they were most influential.**
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, each of which was identified by climwin as the top temperature-related driver *at four of the eight sites*. $T_{min}$ was the top driver at only *one site (BCI)*, and $T_{mean}$ was never identified as the top variable within the temperature group (Fig. `r clim_sens`).
Within the precipitation group, precipitation amount ($PPT$) was identified as the top variable most frequently (n=*6 of 9* sites), but it was not uncommon that it was surpassed by precipitation frequency ($PDF$; n=*3 of 8* sites). 
Optimal time windows often coincided with a site's peak growing season (n= *# of 10* for temperature variables, *# of 10* for precipitation variables), but exceptions were common. 
At the *5 lowest latitude* sites (BCI, HKK, LT, CB, and SCBI), the optimal window for precipitation variables spanned $\ge$ 8 months, ending during the peak growing months of the year of ring formation. 
At the *3 highest latitude sites* (HF, ZOF, and SC), the optimal window for precipitation variables was a short ($\le$ 3 months) window during the previous growing season.
Optimal windows for temperature variables tended to be shorter, the longest being a *6*-month period during the summer (wet season) at HKK. 
At two of the higher-latitude temperate sites (HF and Žofín), temperatures were most influential during late winter/ early spring.
There were also a few instances where previous growing season conditions had the strongest influence.

@@

**Across the three metrics of growth, the "landscape" of climate effects over various time windows was generally similar, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).**
Specifically, $RW$, $BAI$, and $\Delta AGB$ consistently exhibited similar strength of correlation and direction of response to climate variables within the temperature and precipitation variable groups. 
In some cases (n= # of #), both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In # cases, *climwin* identified the same climate variable but different (sometimes overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, in *2 cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at BCI, precipitation variable group at HF; Figs. `r clim_sens`, **S5, S11**)*, *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, we focus on the climate drivers identified when $RW$ was the growth metric and for the full set of cores (i.e., including those for which $DBH$ could not be reconstructed.)

### 2020-11-16 doc/manuscript/MEE_manuscript_SI.Rmd
climwin figures in SI

# MOVE FIGURE FILES
#traditional comparison
flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/formal_comparison/figures/", "^climwin_vs_dcc_.+[.]png$", full.names = TRUE)
file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/doc/manuscript/tables_figures/SI_figures/traditional_comparison/")

@@

#SI figure numbers
traditional_comparison = "S1"

#SI figure captions
traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals."

@@

\newpage 
## Figure S1. Comparison of our approach with traditional methods of identifying climate signals

@@

**Figure S1 | (Comparison of traditional approaches with ours).**  (Example is Cedar Breaks PSME response to (a) PRE and (b) TMX. We have 3 more sites...)

\newpage
## Figure S2. Comparison of climwin output across growth metrics for the temperature variable group at Little Tesuque (New Mexico, USA)
```{r , echo=FALSE, fig.align = "center", out.width='75%', fig.pos='H'}

 # knitr::include_graphics(paste0(path_to_dir, "/results/climwin_plots_combined/NewMexico_pet.png"))
```

**Figure S2 | Comparison of climwin output across growth metrics for the temperature variable group at Little Tesuque (New Mexico, USA).** Here, *climwin* identified potential evapotranspiration ($PET$) as the strongest climate variable across all three metrics of growth ($RW$, BAI, $\Delta AGB$) and regardless of whether all cores were included in the analysis, or only those for which DBH could be reconstructed ($RW$-trees with $DBH$, BAI, $\Delta AGB$).

\newpage
## Figure S3. Comparison of climwin output across growth metrics for the precipitation variable group at Little Tesuque (New Mexico, USA)

```{r , echo=FALSE, fig.align = "center", out.width='75%', fig.pos='H'}

 # knitr::include_graphics(paste0(path_to_dir, "/results/climwin_plots_combined/NewMexico_wet_pre.png"))
```

**Figure S3 | Comparison of climwin output across growth metrics for the precipitation variable group at Little Tesuque (New Mexico, USA).** Here, *climwin* identified precipitation ($PPT$) as the strongest climate variable for $RW$ and $BAI$, but precipitation day frequency ($PDF$) as the strongest climate variable for $\Delta AGB$.

\newpage
## Figure S4. Comparison of climwin output across growth metrics for the precipitation variable group at Harvard Forest (Massachusetts, USA)
![**Figure S4 | Comparison of climwin output across growth metrics for the precipitation variable group at Harvard Forest (Massachusetts, USA).** Here, *climwin* identified precipitation frequency ($PDF$) as the strongest climate variable for $RW$, but precipitation amount ($PPT$) as the strongest climate variable for $BAI$ and $\Delta AGB$. The optimal time window (circled) also differed across growth metrics.](tables_figures/SI_figures/HarvardForest_pre_wet.png)

>>>>

# MOVE FIGURE FILES (doing this because I got an error when using the absolute path to call figures)
#traditional comparison
flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/formal_comparison/figures/", "^climwin_vs_dcc_.+[.]png$", full.names = TRUE)
file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/doc/manuscript/tables_figures/SI_figures/traditional_comparison/")

# climwin plots
flist <- list.files("/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/results/climwin_plots_combined/", full.names = TRUE)
file.copy(flist, "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/EcoClimLab/ForestGEO-climate-sensitivity/doc/manuscript/tables_figures/SI_figures/climwin_plots_combined/")

@@

#SI figure numbers
all_traditional_comparisons = "S1-S4"
SCBI_traditional_comparison = "S1"
Zofin_traditional_comparison = "S2"
CB_traditional_comparison = "S3"
SC_traditional_comparison = "S4"

all_climwin = "S5-S7"
climwin_sameV_sameW = "S5"
climwin_sameV_diffW = "S6"
climwin_diffV_diffW = "S8"

#SI figure captions
SCBI_traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals: LITU at SCBI."
Zofin_traditional_comparison_legend <- "Figure S2. Comparison of our approach with traditional methods of identifying climate signals: ABAL at Zofin."
CB_traditional_comparison_legend <- "Figure S3. Comparison of our approach with traditional methods of identifying climate signals: PSME at Cedar Breaks."
SC_traditional_comparison_legend <- "Figure S4. Comparison of our approach with traditional methods of identifying climate signals: PIMA at Scotty Creek."

climwin_sameV_sameW_legend = "Figure S5. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S6. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S7. (TMP/TMN at BCI)"

@@

\newpage 
## `r SCBI_traditional_comparison_legend`
![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pre.png){ width=70% }

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_SCBI_LITU_pet.png){ width=70% }

**`r SCBI_traditional_comparison_legend`.**  ( NEEDS  FULL CAPTION.. response to (a) PRE and (b) PET.)

## `r Zofin_traditional_comparison_legend`

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_wet.png){ width=70% }

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_Zofin_ABAL_tmx.png){ width=70% }

**`r Zofin_traditional_comparison_legend`.**  ( NEEDS  FULL CAPTION.. response to (a) PDF and (b) TMX )

\newpage 
## `r CB_traditional_comparison_legend`

@@

**`r CB_traditional_comparison_legend`.**  ( NEEDS  FULL CAPTION..  response to (a) PRE and (b) TMX. )

\newpage 
## `r CB_traditional_comparison_legend`

![ ](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_pre.png){ width=70% }

![](tables_figures/SI_figures/traditional_comparison/climwin_vs_dcc_ScottyCreek_PIMA_tmx.png){ width=70% }

**`r CB_traditional_comparison_legend`.**  ( NEEDS  FULL CAPTION..  response to (a) PRE and (b) TMX. )

\newpage
## `r climwin_sameV_sameW_legend`
![](tables_figures/SI_figures/climwin_plots_combined/SCBI_pre.png)

**`r climwin_sameV_sameW_legend`** Here, *climwin* identified **GIVE WINDOW** precipitation ($PPT$) as the strongest climate variable across all four analyses ($RW$ with and without trees for which $DBH$ could not be reconstructed, $BAI$, $\Delta AGB$).

\newpage
## `r climwin_sameV_diffW_legend`
![](tables_figures/SI_figures/climwin_plots_combined/SCBI_pet.png)

**`r climwin_sameV_diffW_legend`** Here, *climwin* identified potential evapotranspiration ($PET$) as the strongest climate variable across all four analyses ($RW$ with and without trees for which $DBH$ could not be reconstructed, $BAI$, $\Delta AGB$), but a different window (**GIVE WINDOW**) was chosen for  $BAI$ and $\Delta AGB$ than for $RW$ (**GIVE WINDOW**).

\newpage
## `r climwin_diffV_diffW_legend`
![](tables_figures/SI_figures/climwin_plots_combined/BCI_tmp_tmn.png)

**`r climwin_diffV_diffW_legend`** Here....
